---
title: "western_blotting"
output: pdf_document
date: "2022-11-16"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Western blotting - data analysis 

Treatments: A - 0ng/ml, B - 10ng/ml, C - 100ng/ml, D - 500ng/ml 

The experiment used three different lots of cells (i.e., 1, 2 and 3) coming from 3 different healthy adults (Lonza). 

```{r}
library(tidyverse)
library(readr) 
library(tidyr) 
library(rstanarm)
library(bayesplot) 
library(tidybayes)
library(emmeans) 
library(scico) 
#setwd
```


```{r}
wb_data = read.csv("optical_density_akt.csv")
head(wb_data)

#summary statistics
group_by(wb_data, treatment) %>%
  summarise(
    count = n(),
    mean = mean(values, na.rm = TRUE),
    sd = sd(values, na.rm = TRUE)) 

wb_data$Lot <- as.character(wb_data$Lot)

#visualize data 
wb_plot <- ggplot(wb_data, aes(x = treatment, y = values)) + 
 geom_jitter(width=0.05, aes(colour = Lot), size = 4) +
  stat_summary(fun='mean',  width=0.4, geom='crossbar') + 
  xlab("Intelectin-1 concentration (ng/ml)") + ylab("Phosphorylated:total Akt (AU)") +
  scale_x_discrete(labels=c("A" = "0", "B" = "10", "C" = "100", "D" = "500")) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 20))

wb_plot
```

#Bayesian ANOVA 

We want to examine whether the treatments had an effect on pAkt:Akt.

```{r}
bayes_aov_priors <- stan_lm(values ~ treatment, 
                            prior_intercept = normal(location = mean(wb_data$values), 
                                                     scale = sd((wb_data$values))), 
                            prior = R2(location = 0.5, what = 'mode'), 
                            prior_PD = TRUE, 
                            data = wb_data, seed = 123)
plot(bayes_aov_priors, plotfun = 'dens')
prior_summary(bayes_aov_priors)

#Linear model
bayes_aov <- stan_lm(values ~ treatment, 
                     prior_intercept = normal(location = mean(wb_data$values),
                                                        scale = sd((wb_data$values))),
                     prior = R2(location = 0.5, what = 'mode'), 
                     data = wb_data, 
                     seed = 123)

# plots
plot(bayes_aov, plotfun = 'dens')

# summary, 90% HDI
summary(bayes_aov, probs = c(0.05, 0.5, 0.95), digits = 3)
```

#### Is there an effect of the categorical predictor?

```{r}
effect_draws <- emmeans(bayes_aov, spec = ~treatment) %>% gather_emmeans_draws()
head(effect_draws)

div_cols <- scico(4, palette = 'roma') # from the scales lib
effect_plt <-  ggplot(effect_draws, aes(x = .value, fill = treatment)) +   
  stat_halfeye(.width = 0.9, point_interval = median_hdi, alpha = 0.7) + 
  scale_fill_manual(values = div_cols, 
                    labels=c('0 ng/ml', '10 ng/ml', '100 ng/ml', '500 ng/ml')) +
  guides(fill = guide_legend(title="ITLN1 treatment")) + 
  theme(axis.text.y=element_blank(), 
        axis.title.y=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 20)) +
  xlab("Phosphorylated:total Akt (AU)")

effect_plt #figure with overalapping posterior distribution
```

```{r}
effect_draws_summary <- emmeans(bayes_aov, spec = ~treatment, level = 0.9)
effect_draws_summary #table with means and90 %HDI
```

##How much variance in the pAKT/AKT ratio estimates does ```treatment``` account for?

```{r}
plot(bayes_aov, plotfun = 'mcmc_areas', pars = 'R2', prob = 0.9)
summary(bayes_aov, pars = 'R2', probs = c(0.05, 0.5, 0.95), digits = 3)
```


##How big are the differences between the different treatments?

```{r}
pairs(emmeans(bayes_aov, spec = ~treatment, level = 0.9))
diffs <- pairs(emmeans(bayes_aov, spec = ~treatment)) %>% gather_emmeans_draws()
ggplot(diffs, aes(x = .value, y = contrast)) + stat_halfeye(.width = 0.9, point_interval = mean_hdi)
```

```{r}
effect_draws_summary <- emmeans(bayes_aov, spec = ~treatment, level = 0.9)
ctrl_uHDI <- effect_draws %>% filter(treatment == 'A') %>% summarise(uHDI = quantile(.value, 0.95)) %>% pull() 
effect_draws %>% group_by(treatment) %>% summarise(prop_gt = sum(.value > ctrl_uHDI)/length(.value))
```

#Posterior predictive checking.

```{r}
div_cols <-scico(n = 4, palette = "roma")

yrep <- add_predicted_draws(bayes_aov, newdata = wb_data, 
                            model = bayes_aov, ndraws = 500)

ggplot(yrep, aes(.prediction, treatment, fill = treatment)) + 
  stat_slab() +  scale_fill_manual( values = div_cols) +
  geom_point(aes(x = values), data = wb_data, alpha = 0.5, 
             position = position_nudge(y = -0.05))
```


```{r}
puro_data = read.csv("optical_density_puro.csv")
head(puro_data)

#summary statistics
group_by(puro_data, treatment) %>%
  summarise(
    count = n(),
    mean = mean(values, na.rm = TRUE),
    sd = sd(values, na.rm = TRUE)) 

puro_data$Lot <- as.character(puro_data$Lot)

#visualize data 
puro_plot <- ggplot(puro_data, aes(x = treatment, y = values)) + 
 geom_jitter(width=0.05, aes(colour = Lot), size = 4) +
  stat_summary(fun='mean',  width=0.4, geom='crossbar') + 
  xlab("Intelectin-1 concentration (ng/ml)") + ylab("Puromycin incorporation (AU)") +
  scale_x_discrete(labels=c("A" = "0", "B" = "10", "C" = "100", "D" = "500")) + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 20))

puro_plot
```

#Bayesian ANOVA 

We want to examine whether the treatments had an effect on puro expression.

```{r}
bayes_aov_priors2 <- stan_lm(values ~ treatment, 
                            prior_intercept = normal(location = mean(puro_data$values), 
                                                     scale = sd((puro_data$values))), 
                            prior = R2(location = 0.5, what = 'mode'), 
                            prior_PD = TRUE, 
                            data = puro_data, seed = 123)
plot(bayes_aov_priors2, plotfun = 'dens')
prior_summary(bayes_aov_priors2)

#Linear model
bayes_aov2 <- stan_lm(values ~ treatment, 
                     prior_intercept = normal(location = mean(puro_data$values),
                                                        scale = sd((puro_data$values))),
                     prior = R2(location = 0.5, what = 'mode'), 
                     data = puro_data, 
                     seed = 123)

# plots
plot(bayes_aov2, plotfun = 'dens')

# summary, 90% HDI
summary(bayes_aov2, probs = c(0.05, 0.5, 0.95), digits = 3)
```


#### Is there an effect of the categorical predictor?


```{r}
effect_draws2 <- emmeans(bayes_aov2, spec = ~treatment) %>% gather_emmeans_draws()
head(effect_draws2)

div_cols2 <- scico(4, palette = 'roma') # from the scales lib
effect_plt2 <-  ggplot(effect_draws2, aes(x = .value, fill = treatment)) +   
  stat_halfeye(.width = 0.9, point_interval = median_hdi, alpha = 0.7) + 
  scale_fill_manual(values = div_cols2, 
                    labels=c('0 ng/ml', '10 ng/ml', '100 ng/ml', '500 ng/ml')) +
  guides(fill = guide_legend(title="ITLN1 treatment")) + 
  theme(axis.text.y=element_blank(), 
        axis.title.y=element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 20)) +
  xlab("Puromycin incorporation (AU)")

effect_plt2 #figure with overalapping posterior distribution
```


```{r}
effect_draws_summary2 <- emmeans(bayes_aov2, spec = ~treatment, level = 0.9)
effect_draws_summary2 #table with means and 90% HDI
```


##How much variance in the puromycin expression does ```treatment``` account for?

```{r}
plot(bayes_aov2, plotfun = 'mcmc_areas', pars = 'R2', prob = 0.9)
summary(bayes_aov2, pars = 'R2', probs = c(0.05, 0.5, 0.95), digits = 3)
```

##How big are the differences between the different treatments?

```{r}
pairs(emmeans(bayes_aov2, spec = ~treatment, level = 0.9))
diffs2 <- pairs(emmeans(bayes_aov2, spec = ~treatment)) %>% gather_emmeans_draws()
ggplot(diffs2, aes(x = .value, y = contrast)) + stat_halfeye(.width = 0.9, point_interval = mean_hdi)
```


```{r}
effect_draws_summary2 <- emmeans(bayes_aov2, spec = ~treatment, level = 0.9)
ctrl_uHDI2 <- effect_draws2 %>% filter(treatment == 'A') %>% summarise(uHDI = quantile(.value, 0.95)) %>% pull() 
effect_draws2 %>% group_by(treatment) %>% summarise(prop_gt = sum(.value > ctrl_uHDI2)/length(.value))
```
